---
title: "General Social Survey"
author: Jaynell Anderson
date: 10/7/2019
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

both variables in this analysis are categorical; So we use bar graphs and tables to summarize the data, and a chi squared test for inference.

$H_0$ ; There is no relationship between gun ownership and whether one is in favor of capital punishment.

$H_A$ ; There is a relationship between gun ownership and whether one is in favor of capital punishment.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~CAPPUN+OWNGUN , data=mydata4, main='Capital Punishment and Gun Ownership', type='percent', xlab='Capital Punishment')
```

This bar chart shows that those who favor captial punishment are equally likely to own a gun, whilst those who oppose capital punishment are less lkely to own a gun.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~CAPPUN+OWNGUN, data=mydata4)
rowPerc(tab1)

```

This table shows that those who favor CAPPUN are more likely to own a gun than those who are opposed to CAPPUN. It also shows that those who are opposed to CAPPUN are less likely to be a gun owner.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```
the p-value for the Chisq test is 0.02 < 0.05 so I reject the null hypothesis and conclude I have evidence to support the contention the gun owner ship and whether on favor capital punishment are related.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is: belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

both variables in this analysis are categorical; So we use bar graphs and tables to summarize the data, and a chi squared test for inference.

$H_0$ ; There is no relationship between belief in the big bang theory and belief in the afterlife..

$H_A$ ; There is a relationship between belief in the big bang theory and belief in the afterlife.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, main= "Belief in Big Bang theory and Postlife", type="percent")
```

This graph shows that there is a very slight difference between the two graphs so there is no real relationship between believing in the big bang theory and believing in postlife.

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~BIGBANG+POSTLIFE , data=mydata4)
rowPerc(tab1)
```

This shows that theres no marked difference in belief in the big bang theory and belief in the postlife
### Inferential Results

```{r}
chisqtestGC(tab1)
```

The P-value for the chisqtest is 0.08 > 0.05 so i fail to reject the null hypothesis and there is no evidence that there is a relationship between belief in the big bang theory and the belief of postlife.
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine: if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

both variables in this analysis are categorical; So we use bar graphs and tables to summarize the data, and a chi squared test for inference.

$H_0$ ; There is no relationship between an individuals opinion on the death penalty and the individuals race.

$H_A$ ; There is a relationship between an individuals opinion on the death penalty and the individuals race.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, main="Opinion of CAPPUN v. the RACE",type="percent")
```

This graph shows that while whites show their opinion of CAPPUN and their races are about the same percentage in favor as oppose CAPPUN. Whereas blacks oppose capital punishment and others tend to favor CAPPUN.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
tab1 = xtabs(~CAPPUN+RACE , data=mydata4)
rowPerc(tab1)
```

This table shows that whites do show a close relationship between their opinion on CAPPUN and their race. Where blacks tend to oppose CAPPUN and Others favor it.
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
chisqtestGC(tab1)
```
The P-value for the chisqtest is 0.0009<0.5 so we reject the null hypothesis. we have evidence that there is a relationship between race and opinion of capital punishment.
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
For question 1- the P-value of the Chisq test is 0.02 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the contention the gun owner ship and whether on favor capital punishment are related.

For question 2- The P-value for the chisqtest is 0.08 > 0.05 so I fail to reject the null hypothesis and there is no evidence that there is a relationship between belief in the big bang theory and the belief of postlife.

For question 3 - The P-value for the chisqtest is 0.0009<0.5 so we reject the null hypothesis. we have evidence that there is a relationship between race and opinion of capital punishment.
